Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO: a Bayesian study

### 2016-04-14 Paper/paper.tex
s5 text first pass done

% \includegraphics[width=1.\columnwidth]{plots/FillBetweenNormErrorBarsLambda_vs_N_L800.pdf}

@@

Posterior probability distributions for $\lambdans$ (colored curves), and 
associated $90\%$ confidence intervals (grey vertical lines), shown for 
different number of accumulated observations $N$.

@@

\includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L2000.pdf}
\includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1500.pdf}\\
\includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1000.pdf}
\includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
\includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L500.pdf}
\caption{Filled-region plots showing the median and $90\%$ confidence intervals
for $\lambdans$ measurement, as a function of the number of observed events $N$.
Different panels correspond to different injected $\lambdans$ values, which is shown
in their title. In each panel are shown two independent filled regions. 
 One in green shows the recovered confidence interval
 for $\lambdans$ (on the left y-axis), normalized by its true value,
 as a function of the 
 number of observed events $N$. The recovered median is shown by the
 green line-circled curve. The pair of horizontal green dashed (dotted)
 lines show the $\pm 50\%$ ($\pm 25\%$) symmetric error bounds around
 the true $\lambdans$ value, which is shown at y-value $=1$ by definition.
 The second filled region, in grey, shows the recovered
 confidence interval for neutron star compactness, without normalizing.
 The corresponding y-values are shown on the *right* y-axis. The horizontal grey
 dashed line traces the true value of $\lambdans$, and is what we expect
 the filled region to approach with increasing number of observations.
 Finally, the grey dotted lines are contours of $1/\sqrt{N}$, drawn to aid the eye.
\label{fig:TT_Lambda_vs_N_L500_2000_CI90_0}

@@

Posterior probability distributions for $\lambdans$ (colored curves), and 
associated $90\%$ confidence intervals (grey vertical lines), shown for 
different number of accumulated observations $N$.

@@

\caption{This figure shows the fractional difference between the median

@@

@@

\label{fig:TT_Lambda_vs_N_L500_2000_CI90_0_AllInOne}

@@

\label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0}

@@

\label{fig:TT_Lambda_vs_Lambda_L500_2000_CI90_0_AllInOne}

@@

\textbf{Multiple identical sources at low SNR: }\label{s2:identical_multiple}
% 
First we show a crude calculation which tells us that approximately $8$ (or $16$) 
observations
of NSBH signals with $\rho\geq 10$ distributed uniformly in effective volume, would
give us similarly accurate a measurement of $\lambdans$ as would a single detection
with $\rho=50$ (or $\rho=70$).

We want to distribute signals in effective distance $\deff$, which is a combination of 
luminosity distance, inclination angle of the binary's orbital angular momentum
with respect to the detector, and its various sky angles. The signal strength 
scales inversely with $\deff$, if it comprises only of the dominant $(2,2)$ mode, allowing
for various angle-dependent factors to scale out beside the luminosity distance.
% 
In order to consider multiple events, lets imagine a population uniformly distributed
in effective volume, i.e. within a sphere of radius $\deff^\mathrm{max}$. The 
radius of this sphere is set by the lowest SNR that is distinguishable from noise
by LIGO searches. Now, divide the sphere into $I$ shells of equal width. The radius
of the $i$-th sphere would then be $D_i = \deff^\mathrm{max} (i - 1/2)/I$. The 
measurement uncertainty in $\lambdans$ scales inversely with the SNR, and hence
directly with the effective distance to the source. Therefore, if we have a 
measurement error $\sigma_0$ for a source located at $\deff = D_0$, the same error
for the same source located within the $i$-th shell would be 
$\sigma_i = \sigma_0 \dfrac{D_i}{D_0}.$
Independent measurements of $\lambdans$ for identical sources at different distances
would have their combined error given by
\begin{equation}\label{eq:1oversigma}
\frac{1}{\sigma^2} = \sum_{i=1}^I \frac{N_i}{\sigma_i^2} = \left(\frac{D_0}{\sigma_0}\right)^2 \sum_{i=1}^I\frac{N_i}{D_i^2},
\end{equation}
where $N_i$ is the number of sources detected in the $i$-th shell.
% , and is a random
% variable, with its probability proportional to the volume of the shell, i.e.
% $$
% p(N_i) \propto \frac{V_i}{V_\mathrm{total}} \propto \dfrac{\left(\deff^\mathrm{max} \frac{i}{I}\right)^3 - \left(\deff^\mathrm{max} \frac{i-1}{I}\right)^3}{(\deff^\mathrm{max})^3} \propto \frac{1}{I^3} [i^3 - (i-1)^3].
% $$
% Therefore, the expected number of detections in the $i$-th shell would be
% $$
% \langle N_i\rangle = \frac{N}{I^3} [i^3 - (i-1)^3],
% $$
% where $N=\sum_{i=1}^I N_i$ is the total number of NSBH detections. $N$ is expected
% to be Poisson distributed around the mean detection rate
% $\mathcal{R}\equiv\langle N\rangle$, where $\mathcal{R}$ can vary from $0.6-1000$ per
% $\mathrm{Gpc}^3$ per year~\cite{Abadie:2010cfa}. If there are $n$ resulting detections
% a year per unit effective volume, then
% \begin{equation}
% \langle N\rangle = \int_0^{\deff^\mathrm{max}} 4\pi n D^2 \D D = \frac{4\pi}{3} n (\deff^\mathrm{max})^3,
% \end{equation}
% and
% \begin{eqnarray}
%  \langle \frac{1}{\sigma^2}\rangle &=& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{\deff^\mathrm{max}} \frac{4\pi n D^2 }{D^2}\D D\\
%  &=& \left(\frac{D_0}{\sigma_0}\right)^2 4\pi n \deff^\mathrm{max},
% \end{eqnarray}
% where in the previous equation we have converted the summation in Eq.~\ref{eq:1oversigma} 
% to an integral. 
Continuing on this line of inquiry, Ref.~\cite{Markakis:2010mp} calculates the 
root-mean-square averaged measurement error from 
$\langle N\rangle:=\langle\sum N_i\rangle$ sources distributed uniformly in
effective volume within a sphere of radius $\deff^\mathrm{max}$ to be
\begin{equation}\label{eq:rmsSigmaIdenticalSources}
 \sigma_{avg} := \frac{1}{\sqrt{\langle1/\sigma^{2}\rangle}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{\sqrt{3\langle N\rangle}},
\end{equation}
given a fiducial pair $(\sigma_0, D_0)$.
Further, it is straightforward to deduce from Eq.~\ref{eq:rmsSigmaIdenticalSources}
that the same measurement certainty as afforded by a single
observation with a high SNR $\rho_c$ can be obtained from $N = \rho_c^2/300$ observations
uniformly distributed in effective volume with SNRs $\geq 10$. E.g., to get to the level 
of certainty afforded at SNR$=70$, we would need $49/3\approx 16-17$ realistic detections.

There are two main caveats to this rudimentary calculation, which was outlined in 
Ref.~\cite{Markakis:2010mp}, (i) it applies to identical sources,
which is astrophysically next to impossible to achieve exactly, but could perhaps happen 
approximately; (ii) the combined average measurement error $\sigma_{avg}$ has been defined
as in the first equality in Eq.~\ref{eq:rmsSigmaIdenticalSources}. To overcome both,
we next perform a Bayesian population analysis.

@@

Our goal is to understand the improvement in our measurement of $\lambdans$
with the number of recorded events. To do so, we simulate populations of $N$
observations each, and quantify what we learn from each successive observation 
using Eq.~\ref{eq:lambdaMultiple}. For each population, we determine how
rapidly our median estimate for $\lambdans$ converges to the true value,
and how rapidly our confidence intervals for the same shrink, with increasing
$N$. In order to generate each population, we first fix
the NS properties: (i) NS mass $\mns=1.35M_\odot$, (ii) NS spin $\chins=0$
and (iii) NS tidal deformability $\lambdans=$ fixed value chosen from
$\{500,800,1000,1500,2000\}$. Each event is generated independently, sampling
the remaining source parameters uniformly from the following ranges: (i)
mass-ratio $q\in[2,5]$, (ii) BH spin $\chibh\in[0, 0.75]$, (iii) effective 
distance from within a spherical volume whose radius is given by the lowest 
detectable SNR (which we take as $\rho=10$). We remind ourselves that
once binary parameters are fixed, there is a one-to-one mapping between
effective distance and SNR. Finally, in order to mitigate computational
cost, we make an additional simplification.
We perform full Bayesian parameter estimation for a set of simulated signals
whose parameters are the vertices of the following regular hypercubic grid
(henceforth ``G''):
$q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
The final population is given by substituting all of the $N$ events, generated
as above, with their respective nearest neighbours on the parameter grid G.

In Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0} we show results for a population
of $N=80$ events. The neutron star deformability is fixed at $\lambdans=800$
for all events. Each curve shows the probability distributions for $\lambdans$
as inferred from $N$ events, with $N$ ranging from $1-50$. We also mark the 
$90\%$ confidence intervals for $\lambdans$ as dark grey vertical lines,
one associated with each probability distribution curve. Distributions are 
normalized to unit area. From the figure, we can visualize the information 
information flow. The first few observations do not have enough information 
to bound $\lambdans$ (with $90\%$ confidence) much more than our prior from
Eq.~\ref{eq:lprior} does. However, the median of these cumulative distributions
track the true value of $\lambdans$ rapidly, reaching close in as few as $4$
observations.

In Fig.~\ref{fig:TT_Lambda_vs_N_L500_2000_CI90_0}, we show our measurement 
uncertainty for $\lambdans$ for different populations, with $\lambdans$ alone
changing between populations. Lets consider the top left panel first, which
corresponds to a population of the most deformable neutron stars considered,
with $\lambdans=2000$. The green filled region shows $\delta\Lambda/\Lambda$
as a function of $N$, where we take $\delta\Lambda$ to be the width of the 
$90\%$ confidence interval on $\lambdans$ after $N$ observations.
%
The green filled circles show the median measured value of $\lambdans$
as a function of $N$. The dashed line at y value $=1$ marks zero error, i.e.
$\delta\Lambda/\Lambda=0$. The dotted and dashed lines around this denote
$\pm 25\%$ and $\pm 50\%$ relative errors, respectively. Independently, the
light-grey filled shown the absolute measurement uncertainty $\delta\Lambda$
on the right y-axis , as a function of $N$. A horizontal grey dashed line
marks the true value of $\lambdans$. Contours of $1/\sqrt{N}$ are shown as 
dotted light grey curves to guide the eye while studying the dependence of
$\delta\Lambda$ on $N$.
%

>>>>

\includegraphics[width=1.\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}

@@

These panels show the accumulation of information from a population whose
$\lambdans=800$.
{\it Left:} Posterior probability distributions for $\lambdans$ (colored 
curves), and associated $90\%$ confidence intervals (grey vertical lines),
shown for different number of accumulated observations $N$. Distributions are
normalized to unit area.
{\it Right:} Measured median value of $\lambdans$ (as solid circles) and the 
associated $90\%$ confidence intervals (as the height of filled region), shown
as a function of number of observations $N$. Solid horizontal line indicates
the true value of $\lambdans$. Dashed and dotted horizontal lines (a pair for each
line-style) demarcate $\pm 25\%$ and $\pm 50\%$ error bounds.

@@

% \includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L2000.pdf}
% \includegraphics[width=.9\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1500.pdf}\\
% \includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L1000.pdf}
% \includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
% \includegraphics[width=.67\columnwidth]{plots/FillBetweenErrorBarsLambda_vs_N_L500.pdf}
% \caption{Filled-region plots showing the median and $90\%$ confidence intervals
% for $\lambdans$ measurement, as a function of the number of observed events $N$.
% Each panel corresponds a unique population with injected $\lambdans$ fixed to
% the value shown in its title. In each panel, the measured median for $\lambdans$
% is shown by filled circles, and the $90\%$ confidence interval is shown by the 
% height of the filled region. In addition, the dashed and 
% dotted lines (2 of each line-style) show the $\pm 25\%$ and $\pm 50\%$ error
% bounds, with respect to the true value.
% \label{fig:TT_Lambda_vs_N_L500_2000_CI90_0}

@@

This filled-region figure shows how our measurement of $\lambdans$ improves with
an increasing number of observed events $N$. Each color corresponds to an
independent populations with the true value of $\lambdans$ fixed as given in
the legend. For each population, we show the measured median (as filled
circles) and $90\%$ confidence intervals for $\lambdans$ measurement (as the
height of the filled region about the median). 

@@

{\it Left:} This figure shows the fractional difference between the median

@@

{\it Right:} 

@@

\label{fig:TT_LambdaMedian_vs_N_AllInOne}

@@

\label{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne}

@@

\label{fig:TT_PowerLawLambdaErrorVsLambda}

@@

\textbf{Multiple identical sources at low SNR: }\label{s2:identical_multiple}
% 
We start with a crude calculation which tells us that approximately $8$ (or $16$) 
observations
of NSBH signals with $\rho\geq 10$ distributed uniformly in effective volume, would
give us similarly accurate a measurement of $\lambdans$ as would a single detection
with $\rho=50$ (or $\rho=70$).

The approximation of using (only) the dominant $l=|m|=2$ multipoles of
gravitational-wave strain to construct the signal, ignoring other $l\neq 2$
multipoles, has the advantage of making various angle parameters 
associated with the source degenerate with its distance. As a result, source
distance, source sky location angles, and the inclination angle of
the binary's orbital angular momentum with respect to the detector,
all combine into a single ``effective distance'' $\deff$ that scales
out as the constant amplitude factor $1/\deff$ (constant because none
of the above parameters vary temporally for aligned-spin binaries).
In order to consider multiple events, lets imagine a population uniformly
distributed in effective volume, i.e. within a sphere of radius 
$\deff^\mathrm{max}$. The radius of this sphere is set by the lowest SNR that 
is distinguishable from noise by LIGO searches. Now, divide the sphere into
$I$ shells of equal thickness. The radius of the $i$-th sphere would then
be $D_i = \deff^\mathrm{max} (i - 1/2)/I$. The measurement uncertainty in
$\lambdans$ scales inversely with SNR, and hence directly with the 
effective distance to the source. Therefore, if we have a measurement error 
$\sigma_0$ for a source located at $\deff = D_0$, the same error for the same
source located within the $i$-th shell would be 
$\sigma_i = \sigma_0 \dfrac{D_i}{D_0}$. The combined error from independent
measurements of $\lambdans$ for identical sources at different distances is
given by
\begin{equation}\label{eq:1oversigma}
\frac{1}{\sigma^2} = \sum_{i=1}^I \frac{N_i}{\sigma_i^2} = \left(\frac{D_0}{\sigma_0}\right)^2 \sum_{i=1}^I\frac{N_i}{D_i^2},
\end{equation}
where $N_i$ is the number of sources detected in the $i$-th shell.
% , and is a random
% variable, with its probability proportional to the volume of the shell, i.e.
% $$
% p(N_i) \propto \frac{V_i}{V_\mathrm{total}} \propto \dfrac{\left(\deff^\mathrm{max} \frac{i}{I}\right)^3 - \left(\deff^\mathrm{max} \frac{i-1}{I}\right)^3}{(\deff^\mathrm{max})^3} \propto \frac{1}{I^3} [i^3 - (i-1)^3].
% $$
% Therefore, the expected number of detections in the $i$-th shell would be
% $$
% \langle N_i\rangle = \frac{N}{I^3} [i^3 - (i-1)^3],
% $$
% where $N=\sum_{i=1}^I N_i$ is the total number of NSBH detections. $N$ is expected
% to be Poisson distributed around the mean detection rate
% $\mathcal{R}\equiv\langle N\rangle$, where $\mathcal{R}$ can vary from $0.6-1000$ per
% $\mathrm{Gpc}^3$ per year~\cite{Abadie:2010cfa}. If there are $n$ resulting detections
% a year per unit effective volume, then
% \begin{equation}
% \langle N\rangle = \int_0^{\deff^\mathrm{max}} 4\pi n D^2 \D D = \frac{4\pi}{3} n (\deff^\mathrm{max})^3,
% \end{equation}
% and
% \begin{eqnarray}
%  \langle \frac{1}{\sigma^2}\rangle &=& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{\deff^\mathrm{max}} \frac{4\pi n D^2 }{D^2}\D D\\
%  &=& \left(\frac{D_0}{\sigma_0}\right)^2 4\pi n \deff^\mathrm{max},
% \end{eqnarray}
% where in the previous equation we have converted the summation in Eq.~\ref{eq:1oversigma} 
% to an integral. 
Continuing on this line of inquiry, Ref.~\cite{Markakis:2010mp} found the 
root-mean-square averaged measurement error from 
$\langle N\rangle:=\langle\sum N_i\rangle$ sources distributed uniformly in
effective volume within a sphere of radius $\deff^\mathrm{max}$ to be
\begin{equation}\label{eq:rmsSigmaIdenticalSources}
 \sigma_{avg} := \frac{1}{\sqrt{\langle1/\sigma^{2}\rangle}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{\sqrt{3\langle N\rangle}},
\end{equation}
given a fiducial pair $(\sigma_0, D_0)$. It is straightforward to deduce from
Eq.~\ref{eq:rmsSigmaIdenticalSources}
that the same measurement certainty as afforded by a single
observation with a high SNR $\rho_c$ can be obtained from $N = \rho_c^2/300$ observations
uniformly distributed in effective volume with SNRs $\geq 10$. E.g., to get to the level 
of certainty afforded at SNR$=70$, we would need $49/3\approx 16-17$ realistic detections.

There are two main caveats to this rudimentary calculation, first outlined in 
Ref.~\cite{Markakis:2010mp}: (i) it applies to identical sources,
which is astrophysically next to impossible to achieve exactly, but could perhaps happen 
approximately; (ii) the combined average measurement error $\sigma_{avg}$ has been defined
as in the first equality in Eq.~\ref{eq:rmsSigmaIdenticalSources}. To overcome both,
we next perform a Bayesian population analysis.

@@

Our goal is to understand the improvement in our measurement of $\lambdans$
with the number of recorded events. To do so, we simulate populations of $N$
observations each, and quantify what we learn from each successive observation 
using Eq.~\ref{eq:lambdaMultiple}. For each population, we determine how
rapidly our median estimate for $\lambdans$ converges to the true value,
and how rapidly our confidence intervals for the same shrink, with increasing
$N$. In order to generate each population, we first fix
the NS properties: (i) NS mass $\mns=1.35M_\odot$, (ii) NS spin $\chins=0$
and (iii) NS tidal deformability $\lambdans=$ fixed value chosen from
$\{500,800,1000,1500,2000\}$. Each event is generated independently, sampling
the remaining source parameters uniformly from the following ranges: (i)
mass-ratio $q\in[2,5]$, (ii) BH spin $\chibh\in[0, 0.75]$, (iii) effective 
distance from within a spherical volume whose radius is given by the lowest 
detectable SNR (which we take as $\rho=10$). We remind ourselves that
once binary parameters are fixed, there is a one-to-one mapping between
effective distance and SNR. Additionally, we consider an alternate population
paradigm (paradigm B, with paradigm A being the standard one described 
above), where populations are generated as above but further restricting
mass-ratios such that source BH masses lie {\it outside} the astrophysical 
mass-gap~\cite{Bailyn:1997xt,Kalogera:1996ci,Kreidberg:2012,Littenberg:2015tpa}.
Finally, in order to mitigate computational cost, we make an additional 
simplification. We perform full Bayesian parameter estimation for a set of 
simulated signals whose parameters are the vertices of the following regular
hypercubic grid (henceforth ``G''):
$q=\{2,3,4,5\}\times\chibh=\{-0.5,0,0.5,0.75\}\times\lambdans=\{500,800,1000,1500,2000\}\times\rho=\{10,20,30,50,70\}$.
The final population is given by substituting all of the $N$ events, generated
as above, with their respective nearest neighbours on the parameter grid G.

In Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0} we show illustrative results
for a single 
population of $N=80$ events. The neutron star deformability is fixed at
$\lambdans=800$ for all events. In the left panel, each curve shows the 
probability distributions for $\lambdans$ as inferred from $N$ events, with $N$
ranging from $1-80$. We also mark the $90\%$ confidence intervals associated
with each of the probability distribution curves. From the figure, we
can visualize information accumulation. The first few observations do not have
enough information to bound $\lambdans$ (with $90\%$ confidence) much more than
our prior from Eq.~\ref{eq:lprior} does. However, the median of these
cumulative distributions appears to track the true value of $\lambdans$ 
rapidly.
% 
In the right panel, we present information derived from the left panel.
The line-circle curve shows the measured median from $N$ observations, with
the latter on the $x$-axis. The pair of dashed (dotted) horizontal lines mark
$\pm25\%$ ($\pm50\%$) error bars. At each $N$, the range spanned by the 
filled region is the $90\%$ confidence interval deduced from the same 
events. This figure somewhat quantifies the qualitative deductions we made
from the left panel. We find that the measured median does track the true value
quickly, reaching $10\%$ accuracy with $10-15$ observations. With the same
information, our confidence intervals also shrink to $\pm 25\%$.
% 
In Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0} we show further results from $4$
independent populations, with the source $\lambdans$ alone changing between
them. The sampled values, i.e.$\lambdans=\{500,1000,1500,2000\}$, 
range from the least to the most deformable neutron stars. As in the 
right panel of Fig.~\ref{fig:TT_Lambda_vs_N_L800_CI90_0}, the line-circle curves
track the median measured $\lambdans$, while the filled regions show
the associated $90\%$ confidence intervals. From the figure, we observe
the following: (i) the shrinkage of confidence interval widths with increasing
$N$ happens in a similar manner for all 
populations, (ii) within $5-10$ observations, median measured $\lambdans$ for
all populations lie within $\pm10\%$ of the true values, and (iii) it takes
approximately $20$ events to distinguish definitively (with $90\%$ confidence)
between deformable NSs with $\lambdans=2000$ and compact NSs with 
$\lambdans=500$, or equivalently to distinguish between hard, moderate and soft
nuclear equations of state. This is comparable to what has been found for
monolithic binaries of neutron stars~\cite{DelPozzo:13,Chatziioannou:2015uea,
Agathos:2015a}.

So far we have discussed individual realizations of NSBH populations. The 
underlying stochasticity of the population generation process makes it
difficult to draw generalized inferences (from a single realization) about the
measurability of $\lambdans$. We mitigate this by drawing multiple
realizations of population generation (i.e. multiple identically constructed
populations). Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} shows the first
results for multiple populations. Lets focus on the left panel first. We
show the median measured value of $\lambdans$ from $100$ independent 
populations with fixed $\lambdans=2000$ as the top set of curves. The one 
highlighted (in green) curve corresponds to the population discussed above.
Similarly, we generate sets of $100$ populations with $\lambdans=500-1500$, and
show the median measured $\lambdans$ as four distinct groups of curves in
Fig.~\ref{fig:TT_LambdaMedian_vs_N_AllInOne} (with the single highlighted
curve in each group corresponding to
Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
Dash-dotted horizontal lines demarcate $\pm50\%$ error intervals around
the true $\lambdans$ values. All of these populations are drawn under
paradigm A (i.e. with BH masses allowed to sample the mass-gap that spans
$2-5M_\odot$).
% 
From the figure, we can conclude with a high likelihood that our measured
$\lambdans$ values will be within $10\%$ of the real value in $15-20$ events
for less deformable neutron stars with $\lambdans\leq 1000$, and in as few as
$10$ events for more deformable stars with $\lambdans\geq 1500$.
% 
The right panel of the figure is identical to the left panel, but drawn under
population paradigm B (no BHs alloewd within the mass-gap). Under this
paradigm, we expectedly find that information accumulation is slower. It would
take $20-40$ detections with $\rho\geq10$ under this paradigm, for our (median)
measured $\lambdans$ to lie within $10\%$ of the true value (fewer for 
more deformable NSs, as one would expect).
% 
This is a promising deduction, as one might expect to see $\mathcal{O}(10)$
events over a few years' timescale with design aLIGO~\cite{Abadie:2010cfa}.

Further, we investigate the statistical uncertainties associated with
$\lambdans$ measurements. We use $90\%$ confidence intervals as our measure of
the same. The left column of 
Fig.~\ref{fig:TT_LambdaError_vs_N_L500_2000_CI90_0_AllInOne} shows results
under paradigm A, where BH mass priors include the mass-gap. In each panel
we show the width of the $90\%$ confidence interval measured from $N$ events
(with $N$ on the $x$-axis), normalized by the true value of $\lambdans$ for
the population. In grey are shown results from each of $100$ population
realizations, with the single population highlighted in color corresponding
to the one we focused on in 
Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}.
The inset shows the same, except that the ordinate is {\it not} normalized by
the true value of $\lambdans$. From top to bottom, the population $\lambdans$
decreases from $\lambdans=2000\rightarrow 500$, corresponding to decreasingly
deformable NSs with softer equations of state. We observe the following:
(i) for moderately-hard to hard equations of state with $\lambdans\geq 1000$,
we can constrain $\lambdans$ within $\pm 50\%$ using only $10-20$ events, and
within $\pm 25\%$ with $25-40$ events; (ii) for softer equations of state with
$\lambdans<1000$, we will achieve the same accuracy with $20-30$ and $50+$ 
events, respectively; and (iii) for the first $5$ or so observations, our
measurement confidence spans the entire prior allowed range 
$\lambdans\in[0,4000]$, as shown by the plateauing towards the left edge in
each panel's inset.
% 
In the right column, each panel panel is identical to its corresponding left
one, with the difference that populations are drawn under paradigm B, which
does {\it not} allow for BH masses to fall within the mass-gap. We find that
for NSs with $\lambdans\leq 1000$, it would take $25-40$ events to
constrain $\lambdans$ within $\pm 50\%$ and $50+$ events to constrain it
within $\pm 25\%$. This is somewhat slower than paradigm A, as is to be
expected since here we preclude the lowest mass-ratios, which correspond to
signals with largest tidal signatures. For $\lambdans>1000$ we find that we
can constrain $\lambdans$ within $\pm 50\%$ with a similar number of events as
for paradigm A, but will need more ($30-40$, as compared to $25-40$) events
to further constrain it to within $\pm 25\%$ of the true value.
% 
These are some of the primary findings of this paper.

Finally, we quantitatively explore the dependence of our statistical
uncertainties for $\lambdans$ on the number of events, as well as on the true
NS deformability itself. First, we will focus on the dependence on $N$. We
assume a power-law dependence of the form
$\delta\lambdans\propto\ 1/N^\alpha$. For each of the $100$ populations 
for each of $\lambdans=500-2000$, we compute the exponent $\alpha$ as a
function of the number of observed events $N$, and show it in 
Fig.~\ref{fig:TT_PowerLawLambdaErrorVsN}. There are $100\times5=500$ curves
on the figure, with one highlighted for each value of population's $\lambdans$.
These highlighted values are only special in the sense that they correspond to
populations discussed earlier in this section (c.f.
Fig.~\ref{fig:TT_Lambda_vs_N_CI90_0}-\ref{fig:TT_LambdaMedian_vs_N_AllInOne}).
We immediately observe two things, (i) there is a globally similar dependence
on $N$ for all populations, and (ii) information accumulates faster than
$1/\sqrt{N}$. In fact, we find that if
$\delta\lambdans\propto\frac{1}{N^\alpha}$, $\alpha$ lines in the range
$0.7_{-0.2}^{+0.2}$.
% 
Next, we focus on the dependence of $\delta\lambdans$ on $\lambdans$ of the
population itself. As suggested by Fisher-matrix studies~\cite{Lackey:2013axa},
and as for $N$, we assume the form $\delta\lambdans\propto\lambdans^\beta$.
From each set of $100$ populations with a given $\lambdans$ value, we draw one
at random, and form a set of $5$ similarly drawn populations, one for each of
$\lambdans=\{500,800,1000,1500,2000\}$. With each set, we determine $\beta$
for different number of observed events $N$. In all, we make $100$ independent
$5-$population sets and show the value of $\beta$ measured from each in 
Fig.~\ref{fig:TT_PowerLawLambdaErrorVsLambda}. We find that the assumed
relation $\delta\lambdans\propto\lambdans^\beta$ gets fairly robust for 
larger values of $N$, with $\beta$ converging to $\beta=0.5^{+0.33}_{-0.33}$.
The fact that $\beta<1$ implies that the relative error
$\delta\lambdans/\lambdans$ decreases with increasing $\lambdans$, while the 
absolute error increases.
% 
From these results, we conclude that the measurement uncertainty for
$\lambdans$ after $N$ observations is
\begin{equation}
 \delta\lambdans\propto \dfrac{\lambdans^{0.5^{+0.33}_{-0.33}}}{N^{0.7_{-0.2}^{+0.2}}}.
\end{equation}
We also find that while these results are inferred from paradigm A populations,
paradigm B gives very similar results.

To summarize, in this section we study the improvement in our measurement of 
NS deformability parameter $\lambdans$ with an increasing number of events. We
do so by simulating plausible populations of disrupting NSBH binaries (with
$\rho\geq 10$). We find that:
(i) for more deformable neutron stars (harder equation of states), the median
measured value of $\lambdans$ comes within $10\%$ of the true value with as 
few as $10$ events, while achieving the same accuracy for softer equations of 
state will take $15-20$ source detections; (ii) the statistical uncertainty
associated with $\lambdans$ measurement shrinks to within $\pm50\%$ with
$10-20$ events, and to within $\pm 25\%$ with $50+$ events, when source 
$\lambdans\geq 1000$; (iii) for softer equations of state, the same could take
$25-40$ and $50+$ events, respectively for the two uncertainty thresholds;
and (iv) if BHs really do observe the astrophysical mass-gap, the information
accumulation is somewhat slower than if they do not. We conclude that within
$20-30$ observations, aLIGO would begin to place very interesting bounds on 
the NS deformability, which would allow us to rule out or rank different
equations of state for neutron star matter.
